---
title: "Matching Procedure"
description: |
  Comparing days with Wind Blowing from the North-East to Other Directions. Adjusting for calendar indicators and other weather variables.
author:
  - name: Léo Zabrocki 
    url: https://lzabrocki.github.io/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/fr/zabrocki-leo/
  - name: Anna Alari 
    url: https://scholar.google.com/citations?user=MiFY320AAAAJ&hl=fr
    affiliation: Sorbonne Université & INSERM
    affiliation_url: https://www.inserm.fr/
  - name: Tarik Benmarhnia
    url: https://profiles.ucsd.edu/tarik.benmarhnia
    affiliation: UCSD & Scripps Institute
    affiliation_url: https://benmarhniaresearch.ucsd.edu/
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
      keep_md: true
      toc: true
      toc_depth: 2
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
# code chunk option
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  layout="l-body-outset",
  dev = "CairoPNG",
  dpi = 600
)
```

In this document, we provide all steps required to reproduce our matching procedure. We compare days where:

* treated units are days where winds blow from the North-East in *t*.
* control units are day winds blow from other directions in *t*.

We adjust for calendar indicators and weather confouding factors.

**Should you have any questions, need help to reproduce the analysis or find coding errors, please do not hesitate to contact me at leo.zabrocki@psemail.eu**

# Required Packages

To reproduce exactly the `script_matching_procedure.html` document, we first need to have installed:

* the [R](https://www.r-project.org/) programming language on your computer 
* [RStudio](https://rstudio.com/), an integrated development environment for R, which will allow you to knit the `script_matching_procedure.Rmd` file and interact with the R code chunks
* the [R Markdown](https://rmarkdown.rstudio.com/) package
* and the [Distill](https://rstudio.github.io/distill/) package which provides the template of this document. 

Once everything is set up, we have to load the following packages:

```{r}
# load required packages
library(knitr) # for creating the R Markdown document
library(here) # for files paths organization
library(tidyverse) # for data manipulation and visualization
library(Rcpp) # for running the matching algorithm
library(optmatch) # for matching pairs
library(igraph) # for pair matching via bipartite maximal weighted matching
```

We load our custom `ggplot2` theme for graphs:

```{r}
# load ggplot custom theme
source(here::here(
  "2.scripts",
  "4.custom_ggplot2_theme",
  "script_theme_tufte.R"
))
# define nice colors
my_blue <- "#0081a7"
my_orange <- "#fb8500"
```

We also have to load the `script_time_series_matching_function.R` located in the **0.script_matching_algorithm** folder and which provides the functions used for matching time series:

```{r, include = TRUE}
# load matching functions
source(
  here::here(
    "2.scripts",
    "3.matching_analysis",
    "0.script_matching_algorithm",
    "script_time_series_matching_function.R"
  )
)
```

# Preparing the Data for Matching
 
### Selecting and Creating Relevant Variables

First, we load the data:

```{r}
# load data
data <-
  readRDS(here::here("1.data", "5.data_for_analysis", "data_for_analysis.RDS")) %>%
  # drop wind direction variable as we use instead wind direction categories
  select(-wind_direction)
```

For each covariate, we create the first daily lags and leads and create a new dataframe called `processed_data`:

```{r}
# create first daily lead for each variable
data_leads <- data %>%
  select(date, mean_no2:wind_direction_categories) %>%
  mutate_at(vars(-date), ~  lead(., n = 1, order_by = date)) %>%
  rename_at(vars(-date), function(x)
    paste0(x, "_lead_", 1))

# create first daily lag for each variable
data_lags <- data %>%
  select(date, mean_no2:wind_direction_categories) %>%
  mutate_at(vars(-date), ~  lag(., n = 1, order_by = date)) %>%
  rename_at(vars(-date), function(x)
    paste0(x, "_lag_", 1))

# create processed_data
processed_data <- left_join(data, data_lags, by = "date") %>%
  left_join(., data_leads, by = "date")
```

We can now define the hypothetical experiment that we would like to investigate.

### Creating Potential Experiments

We defined our potential experiments such that:

* treated units are days where winds blow from the North-East in *t*.
* control units are day winds blow from other directions in *t*.

Below are the required steps to select the corresponding treated and control units whose observations are stored in the matching_data:

```{r}
# construct treatment assignment variable
processed_data <- processed_data %>%
  mutate(is_treated = ifelse(wind_direction_categories == "North-East", TRUE, FALSE))

# remove the days for which assignment is undefined
matching_data = processed_data[!is.na(processed_data$is_treated),]

# susbet treated and control units
treated_units = subset(matching_data, is_treated)
control_units = subset(matching_data,!is_treated)
N_treated = nrow(treated_units)
N_control = nrow(control_units)
```

There are `r N_treated` treated units and  `r N_control` control units. We display the distributionf of treated and control units through time:

```{r, fig.width=10, fig.height=4}
# make stripes graph
graph_stripes_wd_experiment <- matching_data %>%
  mutate(is_treated = ifelse(is_treated == "TRUE", "Treated", "Control")) %>%
  ggplot(., aes(x = date, y = 1, fill = is_treated)) +
  geom_tile() +
  scale_x_date(breaks = scales::pretty_breaks(n = 10)) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_manual(name = "Daily Observations:", values = c(my_blue, my_orange)) +
  xlab("Date") +
  theme_tufte() +
  theme(
    panel.grid.major.y = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank()
  )

# display the graph
graph_stripes_wd_experiment

# save the graph
ggsave(
  graph_stripes_wd_experiment,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_stripes_wd_experiment.pdf"
  ),
  width = 30,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

We save the `matching_data` :

```{r}
# save the matching data
saveRDS(matching_data,
        here::here("1.data", "6.matched_data", "matching_data.Rds"))
```

# Matching Procedure

### Defining Thresholds for Matching Covariates

Below is the code to define the relevant thresholds:

```{r}
# we create the scaling list as it is needed for running the algorithm
# but we do not use it

scaling =  rep(list(1), ncol(matching_data))
names(scaling) = colnames(matching_data)

# instead, we manually defined the threshold for each covariate
thresholds = rep(list(Inf), ncol(matching_data))
names(thresholds) = colnames(matching_data)

# threshold for julian date
thresholds$julian_date = 60

# threshold for weekday
thresholds$weekday = 0

# threshold for holidays
thresholds$holidays_dummy = 0

# threshold for bank days
thresholds$bank_day_dummy = 0

# thresholds for average temperature
thresholds$temperature_average = 5

# threshold for wind speed
thresholds$wind_speed = 1.6

# threshold for wind direction categories lag 1
thresholds$wind_direction_categories_lag_1 = 0

# threshold for rainfall duration
thresholds$rainfall_duration = 0
```

### Running the Matching Procedure

We compute discrepancy matrix and run the matching algorithm:

```{r}
# first we compute the discrepancy matrix
discrepancies = discrepancyMatrix(treated_units, control_units, thresholds, scaling)

# convert matching data to data.frame
matching_data <- as.data.frame(matching_data)

rownames(discrepancies) = format(matching_data$date[which(matching_data$is_treated)], "%Y-%m-%d")
colnames(discrepancies) = format(matching_data$date[which(!matching_data$is_treated)], "%Y-%m-%d")
rownames(matching_data) = matching_data$date

# run the fullmatch algorithm
matched_groups = fullmatch(
  discrepancies,
  data = matching_data,
  remove.unmatchables = TRUE,
  max.controls = 1
)

# get list of matched  treated-control groups
groups_labels = unique(matched_groups[!is.na(matched_groups)])
groups_list = list()
for (i in 1:length(groups_labels)) {
  IDs = names(matched_groups)[(matched_groups == groups_labels[i])]
  groups_list[[i]] = as.Date(IDs[!is.na(IDs)])
}
```

For somes cases, several controls units were matched to a treatment unit. We use the `igraph` package to force pair matching via bipartite maximal weighted matching. Below is the required code:

```{r}
# we build a bipartite graph with one layer of treated nodes, and another layer of control nodes.
# the nodes are labeled by integers from 1 to (N_treated + N_control)
# by convention, the first N_treated nodes correspond to the treated units, and the remaining N_control
# nodes correspond to the control units.

# build pseudo-adjacency matrix: edge if and only if match is admissible
# NB: this matrix is rectangular so it is not per say the adjacendy matrix of the graph
# (for this bipartite graph, the adjacency matrix had four blocks: the upper-left block of size
# N_treated by N_treated filled with 0's, bottom-right block of size N_control by N_control filled with 0's,
# top-right block of size N_treated by N_control corresponding to adj defined below, and bottom-left block
# of size N_control by N_treated corresponding to the transpose of adj)
adj = (discrepancies < Inf)

# extract endpoints of edges
edges_mat = which(adj, arr.ind = TRUE)

# build weights, listed in the same order as the edges (we use a decreasing function x --> 1/(1+x) to
# have weights inversely proportional to the discrepancies, since maximum.bipartite.matching
# maximizes the total weight and we want to minimize the discrepancy)
weights = 1 / (1 + sapply(1:nrow(edges_mat), function(i)
  discrepancies[edges_mat[i, 1], edges_mat[i, 2]]))

# format list of edges (encoded as a vector resulting from concatenating the end points of each edge)
# i.e c(edge1_endpoint1, edge1_endpoint2, edge2_endpoint1, edge2_endpoint1, edge3_endpoint1, etc...)
edges_mat[, "col"] = edges_mat[, "col"] + N_treated
edges_vector = c(t(edges_mat))

# NB: by convention, the first N_treated nodes correspond to the treated units, and the remaining N_control
# nodes correspond to the control units (hence the "+ N_treated" to shift the labels of the control nodes)

# build the graph from the list of edges
BG = make_bipartite_graph(c(rep(TRUE, N_treated), rep(FALSE, N_control)), edges = edges_vector)

# find the maximal weighted matching
MBM = maximum.bipartite.matching(BG, weights = weights)

# list the dates of the matched pairs
pairs_list = list()
N_matched = 0
for (i in 1:N_treated) {
  if (!is.na(MBM$matching[i])) {
    N_matched = N_matched + 1
    pairs_list[[N_matched]] = c(treated_units$date[i], control_units$date[MBM$matching[i] -
                                                                            N_treated])
  }
}

# transform the list of matched pairs to a dataframe
matched_pairs <- enframe(pairs_list) %>%
  unnest(cols = "value") %>%
  rename(pair_number = name,
         date = value)
```

The hypothetical experiment we set up had `r N_treated` treated units and `r N_control` control units. The matching procedure results in `r N_matched` matched treated units.

We finally merge the `matched_pairs` with the `matching_matching_data` to retrieve covariates values for the matched pairs and save the data:

```{r}
# select the matched data for the analysis
final_data <- left_join(matched_pairs, matching_data, by = "date")

# save the matched data
saveRDS(final_data,
        here::here("1.data", "6.matched_data", "matched_data.Rds"))
```





