---
title: "Comparing the Matched Data to the Initial Data"
description: |
  Comparing days with Wind Blowing from the North-East to Other Directions. Adjusting for calendar indicators and other weather variables.
author:
  - name: Léo Zabrocki 
    url: https://lzabrocki.github.io/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/fr/zabrocki-leo/
  - name: Anna Alari 
    url: https://scholar.google.com/citations?user=MiFY320AAAAJ&hl=fr
    affiliation: Sorbonne Université & INSERM
    affiliation_url: https://www.inserm.fr/
  - name: Tarik Benmarhnia
    url: https://profiles.ucsd.edu/tarik.benmarhnia
    affiliation: UCSD & Scripps Institute
    affiliation_url: https://benmarhniaresearch.ucsd.edu/
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
      keep_md: true
      toc: true
      toc_depth: 2
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
# code chunk option
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  layout="l-body-outset",
  dev = "CairoPNG",
  dpi = 300
)
```

In this document, we take great care providing all steps and R codes required to compare the matched data to the initial data. We compare days where:

* treated units are days where winds blow from the North-East in *t*.
* control units are day winds blow from other directions in *t*.

We adjust for calendar indicators and weather confouding factors.

**Should you have any questions, need help to reproduce the analysis or find coding errors, please do not hesitate to contact me at leo.zabrocki@psemail.eu**

# Required Packages

To reproduce exactly the `2_script_comparing_two_datasets.html` document, we first need to have installed:

* the [R](https://www.r-project.org/) programming language 
* [RStudio](https://rstudio.com/), an integrated development environment for R, which will allow you to knit the `2_script_comparing_two_datasets.Rmd` file and interact with the R code chunks
* the [R Markdown](https://rmarkdown.rstudio.com/) package
* and the [Distill](https://rstudio.github.io/distill/) package which provides the template for this document. 

Once everything is set up, we have to load the following packages:

```{r}
# load required packages
library(knitr) # for creating the R Markdown document
library(here) # for files paths organization
library(tidyverse) # for data manipulation and visualization
library(ggridges) # for ridge density plots
library(kableExtra) # for table formatting
library(Cairo) # for printing customed police of graphs
library(patchwork) # combining plots
```

We finally load our custom `ggplot2` theme for graphs:

```{r}
# load ggplot custom theme
source(here::here(
  "2.scripts",
  "4.custom_ggplot2_theme",
  "script_theme_tufte.R"
))
# define nice colors
my_blue <- "#0081a7"
my_orange <- "#fb8500"
```

# Comparing Distribution of Covariates in Matched and Initial Datasets

We explore the characteristics of the matched data by comparing the distribution of its covariates to those of the matching data. We load the two datasets and bind them in the `data_all` object:

```{r}
# load matching data
data_matching <-
  readRDS(here::here("1.data", "5.matched_data", "matching_data.Rds")) %>%
  mutate(dataset = "Initial Data")

# load matched data
data_matched <-
  readRDS(here::here("1.data", "5.matched_data", "matched_data.Rds")) %>%
  mutate(dataset = "Matched Data")

# bind the three datasets
data_all <- bind_rows(data_matching, data_matched)
```

### Weather Covariates

We plot below the density distributions of continuous weather covariates for the two datasets:

```{r, fig.width=10, fig.height=5, dev = "CairoPNG"}
# we select continuous weather variables and store them in a long dataframe
data_continuous_weather_variables <- data_all %>%
  select(temperature_average, wind_speed, humidity_average, dataset) %>%
  pivot_longer(
    .,
    cols =  c(temperature_average:humidity_average),
    names_to = "variable",
    values_to = "values"
  ) %>%
  mutate(
    variable =  factor(
      variable,
      levels = c("temperature_average", "humidity_average", "wind_speed")
    ) %>%
      fct_recode(
        .,
        "Temperature Average (°C)" = "temperature_average",
        "Average Humidity (%)" = "humidity_average",
        "Wind Speed (m/s)" = "wind_speed"
      )
  ) %>%
  mutate(dataset = fct_relevel(dataset, "Initial Data", "Matched Data"))

# we plot the density distributions
graph_density_continuous_weather_variables <-
  ggplot(data_continuous_weather_variables,
         aes(
           x = values,
           y = fct_rev(dataset),
           fill = fct_rev(dataset)
         )) +
  geom_density_ridges(colour = NA) +
  scale_fill_manual(values = c(my_blue, my_orange),
                    guide = guide_legend(reverse = TRUE)) +
  xlab("Covariate Value") + ylab("") +
  labs(fill = "Dataset:") +
  facet_wrap( ~ variable, scale = "free_x", ncol = 3) +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    axis.text.x = element_text(
      family = "Roboto Condensed",
      color = "gray18",
      size = 8,
      margin = ggplot2::margin(t = 0, unit = "cm")
    )
  )

# print the graph
graph_density_continuous_weather_variables
````

We plot the proportion of weather categorical variables for the two datasets 

```{r, fig.width=12, fig.height=10, dev = "CairoPNG"}
# we select categorical weather variables and store them in a long dataframe
data_categorical_weather_variables <- data_all %>%
  # select relevant variables
  select(wind_direction_categories, rainfall_duration, dataset) %>%
  drop_na() %>%
  mutate(rainfall_duration = paste("Rainfall Duration:", rainfall_duration, sep = " "))  %>%
  mutate(
    rainfall_duration = case_when(
      rainfall_duration == "Rainfall Duration:   0" ~ "Rainfall Duration: 0 min",
      rainfall_duration == "Rainfall Duration: [  1,  13)" ~ "Rainfall Duration: [1, 13) min",
      rainfall_duration == "Rainfall Duration: [ 13, 159)" ~ "Rainfall Duration: [13, 159) min",
      rainfall_duration == "Rainfall Duration: [159,1440]" ~ "Rainfall Duration: [159, 1440] min"
    )
  ) %>%
  mutate(
    wind_direction_categories = fct_recode(
      wind_direction_categories,
      "Wind Direction: North-East" = "North-East",
      "Wind Direction: South-East" = "South-East",
      "Wind Direction: South-West" = "South-West",
      "Wind Direction: North-West" = "North-West"
    )
  ) %>%
  # transform variables to character
  mutate_all( ~ as.character(.)) %>%
  # transform the data to long to compute the proportion of observations for each variable
  pivot_longer(cols = -c(dataset),
               names_to = "variable",
               values_to = "values") %>%
  # group by dataset, variable and values
  group_by(dataset, variable, values) %>%
  # compute the number of observations
  summarise(n = n()) %>%
  # compute the proportion
  mutate(freq = round(n / sum(n) * 100, 0)) %>%
  # reorder labels of the dataset variable
  ungroup() %>%
  mutate(dataset = fct_relevel(dataset, "Initial Data", "Matched Data"))

# we plot the cleveland dots plots
graph_categorical_weather_variables <-
  ggplot(data_categorical_weather_variables,
         aes(
           x = freq,
           y = fct_rev(dataset),
           fill =  fct_rev(dataset)
         )) +
  geom_segment(aes(
    x = 0,
    xend = freq,
    y =  fct_rev(dataset),
    yend =  fct_rev(dataset)
  )) +
  geom_point(shape = 21,
             color = "black",
             size = 4) +
  scale_fill_manual(values = c(my_blue, my_orange),
                    guide = guide_legend(reverse = TRUE)) +
  facet_wrap( ~ values, scale = "free_x", ncol = 3) +
  ylab("") +
  xlab("Proportion (%)") +
  labs(fill = "Dataset:") +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal",
    axis.text.x = element_text(
      family = "Roboto Condensed",
      color = "gray18",
      size = 8,
      margin = ggplot2::margin(t = 0, unit = "cm")
    )
  )

# print the graph
graph_categorical_weather_variables
````

We combine the `graph_density_continuous_weather_variables` and `graph_categorical_weather_variables` :

```{r, fig.width=28, fig.height=18, dev = "CairoPNG"}
# combine plots
graph_weather_three_datasets <- graph_density_continuous_weather_variables / graph_categorical_weather_variables +
  plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size = 20, face = "bold"))

# save the plot
ggsave(graph_weather_three_datasets, filename = here::here("3.outputs", "2.matching_analysis", "graph_weather_two_datasets.pdf"), 
       width = 35, height = 20, units = "cm", device = cairo_pdf)
````


### Calendar Indicators

We plot the proportions of observations belonging to each day of the week by dataset:

```{r, fig.width=10, fig.height=5, dev = "CairoPNG"}
# compute the proportions of observations belonging to each day of the week by dataset
data_weekday <- data_all %>%
  mutate(weekday = lubridate::wday(date, abbr = FALSE, label = TRUE)) %>%
  select(weekday, dataset) %>%
  mutate(
    weekday = fct_relevel(
      weekday,
      "Monday",
      "Tuesday",
      "Wednesday",
      "Thursday",
      "Friday",
      "Saturday",
      "Sunday"
    )
  ) %>%
  pivot_longer(.,-dataset) %>%
  group_by(name, dataset, value) %>%
  summarise(n = n()) %>%
  mutate(proportion = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(dataset = fct_relevel(dataset, "Initial Data", "Matched Data"))

# we plot the data using cleveland dot plots
graph_weekday <-
  ggplot(data_weekday,
         aes(
           x = as.factor(value),
           y = proportion,
           colour = dataset,
           group = dataset
         )) +
  geom_line(size = 1) +
  scale_colour_manual(values = c(my_orange, my_blue),
                      guide = guide_legend(reverse = FALSE)) +
  ggtitle("Day of the Week") +
  ylab("Proportion (%)") +
  xlab("") +
  labs(colour = "Dataset:") +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal"
  )

# we print the graph
graph_weekday
````

We plot the proportions of observations belonging to bank days and holidays by dataset:

```{r, fig.width=10, fig.height=5, dev = "CairoPNG"}
# compute the proportions of observations belonging to bank days and holidays by dataset
data_bank_holidays <- data_all %>%
  select(bank_day_dummy, holidays_dummy, dataset) %>%
  pivot_longer(.,-dataset) %>%
  mutate(name = recode(name, bank_day_dummy = "Bank Day", holidays_dummy = "Holidays")) %>%
  group_by(name, dataset, value) %>%
  summarise(n = n()) %>%
  mutate(proportion = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(dataset = fct_relevel(dataset, "Data", "Initial Data", "Matched Data")) %>%
  filter(value == 1) %>%
  mutate(name = paste(name, ": True", sep = ""))

# we plot the data using cleveland dot plots
graph_bank_holidays <-
  ggplot(data_bank_holidays,
         aes(
           x = proportion,
           y = as.factor(dataset),
           fill = dataset
         )) +
  geom_segment(aes(
    x = 0,
    xend = proportion,
    y =  fct_rev(dataset),
    yend =  fct_rev(dataset)
  )) +
  geom_point(shape = 21,
             colour = "black",
             size = 4) +
  scale_fill_manual(values = c(my_orange, my_blue),
                    guide = guide_legend(reverse = FALSE)) +
  facet_wrap( ~ name) +
  ggtitle("Bank Days and Holidays") +
  ylab("Proportion (%)") +
  xlab("") +
  labs(fill = "Dataset:") +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal"
  )

# we print the graph
graph_bank_holidays
````

We plot the proportions of observations belonging to each month by dataset:

```{r, fig.width=11, fig.height=5, dev = "CairoPNG"}
# compute the proportions of observations belonging to each month by dataset
data_month <- data_all %>%
  select(month, dataset) %>%
  mutate(
    month = recode(
      month,
      `1` = "January",
      `2` = "February",
      `3` = "March",
      `4` = "April",
      `5` = "May",
      `6` = "June",
      `7` = "July",
      `8` = "August",
      `9` = "September",
      `10` = "October",
      `11` = "November",
      `12` = "December"
    ) %>%
      fct_relevel(
        .,
        "January",
        "February",
        "March",
        "April",
        "May",
        "June",
        "July",
        "August",
        "September",
        "October",
        "November",
        "December"
      )
  ) %>%
  pivot_longer(.,-dataset) %>%
  group_by(name, dataset, value) %>%
  summarise(n = n()) %>%
  mutate(proportion = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(dataset = fct_relevel(dataset, "Initial Data", "Matched Data"))


# we plot the data using cleveland dot plots
graph_month <-
  ggplot(data_month,
         aes(
           x = as.factor(value),
           y = proportion,
           colour = dataset,
           group = dataset
         )) +
  geom_line(size = 1) +
  scale_colour_manual(values = c(my_orange, my_blue),
                      guide = guide_legend(reverse = FALSE)) +
  ggtitle("Month") +
  ylab("Proportion (%)") +
  xlab("") +
  labs(colour = "Dataset:") +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal"
  )

# we print the graph
graph_month
````

We plot the proportions of observations belonging to each year by dataset:

```{r, fig.width=10, fig.height=5, dev = "CairoPNG"}
# compute the proportions of observations belonging to each year by dataset
data_year <- data_all %>%
  select(year, dataset) %>%
  pivot_longer(.,-dataset) %>%
  group_by(name, dataset, value) %>%
  summarise(n = n()) %>%
  mutate(proportion = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(dataset = fct_relevel(dataset, "Initial Data", "Matched Data"))


# we plot the data using cleveland dot plots
graph_year <-
  ggplot(data_year,
         aes(
           x = as.factor(value),
           y = proportion,
           colour = dataset,
           group = dataset
         )) +
  geom_line(size = 1) +
  scale_colour_manual(values = c(my_orange, my_blue),
                      guide = guide_legend(reverse = FALSE)) +
  ggtitle("Year") +
  ylab("Proportion (%)") +
  xlab("") +
  labs(colour = "Dataset:") +
  theme_tufte() +
  theme(
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "horizontal"
  )

# we print the graph
graph_year
````

We combine all plots for calendar variables:

```{r, fig.width=28, fig.height=18, dev = "CairoPNG"}
# combine plots
graph_calendar_three_datasets <-
  (graph_weekday + graph_bank_holidays) / (graph_month + graph_year) +
  plot_annotation(tag_levels = 'A') &
  theme(plot.tag = element_text(size = 20, face = "bold"))

# save the plot
ggsave(
  graph_calendar_three_datasets,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_calendar_two_datasets.pdf"
  ),
  width = 40,
  height = 20,
  units = "cm",
  device = cairo_pdf
)
````