---
title: "Checking Balance Improvement"
description: |
  Comparing days with Wind Blowing from the North-East to Other Directions. Adjusting for calendar indicators and other weather variables.
author:
  - name: Léo Zabrocki 
    url: https://lzabrocki.github.io/
    affiliation: Paris School of Economics
    affiliation_url: https://www.parisschoolofeconomics.eu/fr/zabrocki-leo/
  - name: Anna Alari 
    url: https://scholar.google.com/citations?user=MiFY320AAAAJ&hl=fr
    affiliation: Sorbonne Université & INSERM
    affiliation_url: https://www.inserm.fr/
  - name: Tarik Benmarhnia
    url: https://profiles.ucsd.edu/tarik.benmarhnia
    affiliation: UCSD & Scripps Institute
    affiliation_url: https://benmarhniaresearch.ucsd.edu/
date: "`r Sys.Date()`"
output: 
    distill::distill_article:
      keep_md: true
      toc: true
      toc_depth: 2
---

<style>
body {
text-align: justify}
</style>

```{r setup, include=FALSE}
# code chunk option
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  layout = "l-body-outset",
  dev = "CairoPNG",
  dpi = 600
)
```

In this document, we take great care providing all steps and R codes required to check whether our matching procedure allowed to improve covariates balance. We compare days where:

* treated units are days where winds blow from the North-East in *t*.
* control units are day winds blow from other directions in *t*.

We adjust for calendar indicators and weather confouding factors.

**Should you have any questions, need help to reproduce the analysis or find coding errors, please do not hesitate to contact me at leo.zabrocki@psemail.eu**

# Required Packages

To reproduce exactly the `4_script_checking_balance_improvement.html` document, we first need to have installed:

* the [R](https://www.r-project.org/) programming language 
* [RStudio](https://rstudio.com/), an integrated development environment for R, which will allow you to knit the `4_script_checking_balance_improvement.Rmd` file and interact with the R code chunks
* the [R Markdown](https://rmarkdown.rstudio.com/) package
* and the [Distill](https://rstudio.github.io/distill/) package which provides the template for this document. 

Once everything is set up, we have to load the following packages:

```{r}
# load required packages
library(knitr) # for creating the R Markdown document
library(here) # for files paths organization
library(tidyverse) # for data manipulation and visualization
library(ggridges) # for ridge density plots
library(Cairo) # for printing custom police of graphs
library(patchwork) # combining plots
```

We finally load our custom `ggplot2` theme for graphs:

```{r}
# load ggplot custom theme
source(here::here(
  "2.scripts",
  "4.custom_ggplot2_theme",
  "script_theme_tufte.R"
))
# define nice colors
my_blue <- "#0081a7"
my_orange <- "#fb8500"
````

# Preparing the Data

We load the initial and matched data and bind them together:

```{r}
# load matching data
data_matching <-
  readRDS(here::here("1.data", "5.matched_data", "matching_data.rds")) %>%
  mutate(dataset = "Initial Data")

# load matched data
data_matched <-
  readRDS(here::here("1.data", "5.matched_data", "matched_data.rds")) %>%
  mutate(dataset = "Matched Data")

# bind the two datasets
data <- bind_rows(data_matching, data_matched)
```

We change labels of the `is_treated` variable :

```{r}
data <- data %>%
  mutate(is_treated = ifelse(is_treated == "TRUE", "True", "False"))
```

# Love Plots

### Continuous Weather Covariates

```{r, fig.width=10, fig.height=5}
# compute figures for the love plot
data_weather_continuous <- data %>%
  select(
    dataset,
    is_treated,
    contains("temperature"),
    contains("humidity"),
    contains("wind_speed")
  ) %>%
  pivot_longer(
    cols = -c(is_treated, dataset),
    names_to = "variable",
    values_to = "values"
  ) %>%
  mutate(
    weather_variable = NA %>%
      ifelse(
        str_detect(variable, "temperature_average"),
        "Average Temperature",
        .
      ) %>%
      ifelse(
        str_detect(variable, "humidity_average"),
        "Humidity Average",
        .
      ) %>%
      ifelse(str_detect(variable, "wind_speed"), "Wind Speed", .)
  ) %>%
  mutate(time = "0" %>%
           ifelse(str_detect(variable, "lag_1"), "-1", .) %>%
           ifelse(str_detect(variable, "lead_1"), "+1", .)) %>%
  filter(time != "+1") %>%
  mutate(time = fct_relevel(time, "-1", "0")) %>%
  select(dataset, is_treated, weather_variable, time, values)

data_abs_difference_continuous_weather <-
  data_weather_continuous %>%
  group_by(dataset, weather_variable, time, is_treated) %>%
  summarise(mean_values = mean(values, na.rm = TRUE)) %>%
  summarise(abs_difference = abs(mean_values[2] - mean_values[1]))

data_sd_weather_continuous <-  data_weather_continuous %>%
  filter(is_treated == "True") %>%
  group_by(dataset, weather_variable, time, is_treated) %>%
  summarise(sd_treatment = sd(values, na.rm = TRUE)) %>%
  ungroup() %>%
  select(dataset, weather_variable, time, sd_treatment)

data_love_continuous_weather <-
  left_join(
    data_abs_difference_continuous_weather,
    data_sd_weather_continuous,
    by = c("dataset", "weather_variable", "time")
  ) %>%
  mutate(standardized_difference = abs_difference / sd_treatment) %>%
  select(-c(abs_difference, sd_treatment))

# make the graph
graph_love_plot_continuous_weather <-
  ggplot(
    data_love_continuous_weather,
    aes(
      y = time,
      x = standardized_difference,
      colour = fct_rev(dataset),
      shape = fct_rev(dataset)
    )
  ) +
  geom_vline(xintercept = 0, size = 0.3) +
  geom_vline(xintercept = 0.1,
             color = "black",
             linetype = "dashed") +
  geom_point(size = 4, alpha = 0.8) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  facet_wrap( ~ weather_variable, scales = "free_y") +
  xlab("Standardized Mean Differences") +
  ylab("Day") +
  theme_tufte()

# plot the graph
graph_love_plot_continuous_weather

# save the graph
ggsave(
  graph_love_plot_continuous_weather,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_love_plot_continuous_weather.pdf"
  ),
  width = 30,
  height = 12,
  units = "cm",
  device = cairo_pdf
)
```

### Categorical Weather Covariates

```{r, fig.width=10, fig.height=5}
# compute figures for the love plot
data_weather_categorical <- data %>%
  select(
    dataset,
    is_treated,
    contains("rainfall_duration"),
    "wind_direction_categories_lag_1"
  ) %>%
  drop_na() %>%
  mutate_all( ~ as.character(.)) %>%
  pivot_longer(
    cols = -c(dataset, is_treated),
    names_to = "variable",
    values_to = "values"
  ) %>%
  # group by is_treated, variable and values
  group_by(dataset, is_treated, variable, values) %>%
  # compute the number of observations
  summarise(n = n()) %>%
  # compute the proportion
  mutate(freq = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(
    weather_variable = NA %>%
      ifelse(str_detect(variable, "wind"), "Wind Direction", .) %>%
      ifelse(str_detect(variable, "rainfall"), "Rainfall Duration", .)
  ) %>%
  mutate(time = "t" %>%
           ifelse(str_detect(variable, "lag_1"), "t-1", .) %>%
           ifelse(str_detect(variable, "lead_1"), "t+1", .)) %>%
  filter(time != "t+1") %>%
  mutate(variable = paste(weather_variable, time, sep = " ")) %>%
  select(dataset, is_treated, weather_variable, variable, values, freq) %>%
  pivot_wider(names_from = is_treated, values_from = freq) %>%
  mutate(abs_difference = abs(`True` - `False`)) %>%
  filter(weather_variable != "Wind Direction t")


# create the figure for wind direction
graph_love_plot_wind_direction <- data_weather_categorical %>%
  filter(weather_variable == "Wind Direction") %>%
  ggplot(.,
         aes(
           y = fct_rev(values),
           x = abs_difference,
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0, size = 0.3) +
  geom_point(size = 4, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  facet_wrap( ~ variable, scales = "free_y", ncol = 3) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() +
  theme(axis.text.x = element_text(margin = ggplot2::margin(t = 0, unit = "cm")))

# print the figure for wind direction
graph_love_plot_wind_direction

# save the figure for wind direction
ggsave(
  graph_love_plot_wind_direction,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_love_plot_wind_direction.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)

# create the figure for rainfall dummy
graph_love_plot_rainfall <- data_weather_categorical %>%
  filter(weather_variable == "Rainfall Duration") %>%
  mutate(variable = fct_relevel(variable, "Rainfall Duration t-1", "Rainfall Duration t")) %>%
  ggplot(.,
         aes(
           y = fct_rev(values),
           x = abs_difference,
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0, size = 0.3) +
  geom_point(size = 4, alpha = 0.8) +
  facet_wrap( ~ variable) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() +
  theme(axis.text.x = element_text(margin = ggplot2::margin(t = 0, unit = "cm")))

# print the figure for rainfall dummy
graph_love_plot_rainfall

# save the figure for rainfall dummy
ggsave(
  graph_love_plot_rainfall,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_love_plot_rainfall.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```

### Pollutants

```{r, fig.width=10, fig.height=5}
# compute figures for the love plot
data_pollutants <- data %>%
  select(
    dataset,
    is_treated,
    mean_no2_lag_1,
    mean_o3_lag_1,
    mean_pm10_lag_1,
    mean_pm25_lag_1
  ) %>%
  pivot_longer(
    cols = -c(dataset, is_treated),
    names_to = "variable",
    values_to = "values"
  ) %>%
  mutate(
    pollutant = NA %>%
      ifelse(str_detect(variable, "no2"), "NO2 in t-1", .) %>%
      ifelse(str_detect(variable, "o3"), "O3 in t-1", .) %>%
      ifelse(str_detect(variable, "pm10"), "PM10 in t-1", .) %>%
      ifelse(str_detect(variable, "pm25"), "PM2.5 in t-1", .)
  ) %>%
  select(dataset, is_treated, pollutant, values)

data_abs_difference_pollutants <- data_pollutants %>%
  group_by(dataset, pollutant, is_treated) %>%
  summarise(mean_values = mean(values, na.rm = TRUE)) %>%
  summarise(abs_difference = abs(mean_values[2] - mean_values[1]))

data_sd_pollutants <-  data_pollutants %>%
  filter(is_treated == "True") %>%
  group_by(dataset, pollutant, is_treated) %>%
  summarise(sd_treatment = sd(values, na.rm = TRUE)) %>%
  ungroup() %>%
  select(dataset, pollutant, sd_treatment)

data_love_pollutants <-
  left_join(data_abs_difference_pollutants,
            data_sd_pollutants,
            by = c("dataset", "pollutant")) %>%
  mutate(standardized_difference = abs_difference / sd_treatment) %>%
  select(-c(abs_difference, sd_treatment))

# create the graph
graph_love_plot_pollutants <-
  ggplot(
    data_love_pollutants,
    aes(
      y = fct_rev(pollutant),
      x = standardized_difference,
      colour = fct_rev(dataset),
      shape = fct_rev(dataset)
    )
  ) +
  geom_vline(xintercept = 0, size = 0.3) +
  geom_vline(xintercept = 0.1,
             color = "black",
             linetype = "dashed") +
  geom_point(size = 4, alpha = 0.8) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  xlab("Standardized Mean Differences") +
  ylab("") +
  theme_tufte() +
  theme(axis.text.x = element_text(margin = ggplot2::margin(t = 0, unit = "cm")))

# print the graph
graph_love_plot_pollutants

# save the graph
ggsave(
  graph_love_plot_pollutants,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_love_plot_pollutants.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```


### Calendar Indicators

Create the relevant data:

```{r, fig.width=26, fig.height=15}
# compute figures for the love plot
data_calendar <- data %>%
  mutate(weekday = lubridate::wday(date, abbr = FALSE, label = TRUE)) %>%
  select(dataset,
         is_treated,
         weekday,
         holidays_dummy,
         bank_day_dummy,
         month,
         year) %>%
  mutate_at(vars(holidays_dummy, bank_day_dummy),
            ~ ifelse(. == 1, "True", "False")) %>%
  mutate_all( ~ as.character(.)) %>%
  pivot_longer(
    cols = -c(dataset, is_treated),
    names_to = "variable",
    values_to = "values"
  ) %>%
  # group by is_treated, variable and values
  group_by(dataset, is_treated, variable, values) %>%
  # compute the number of observations
  summarise(n = n()) %>%
  # compute the proportion
  mutate(freq = round(n / sum(n) * 100, 0)) %>%
  ungroup() %>%
  mutate(
    calendar_variable = NA %>%
      ifelse(str_detect(variable, "weekday"), "Day of the Week", .) %>%
      ifelse(str_detect(variable, "holidays_dummy"), "Holidays", .) %>%
      ifelse(str_detect(variable, "bank_day_dummy"), "Bank Day", .) %>%
      ifelse(str_detect(variable, "month"), "Month", .) %>%
      ifelse(str_detect(variable, "year"), "Year", .)
  ) %>%
  select(dataset, is_treated, calendar_variable, values, freq) %>%
  pivot_wider(names_from = is_treated, values_from = freq) %>%
  mutate(abs_difference = abs(`True` - `False`)) %>%
  filter(values != "False")
```

Plot for bank days and holidays:

```{r, fig.width=10, fig.height=5}
# graph for bank days and holidays
graph_love_plot_bank_holidays <- data_calendar %>%
  filter(calendar_variable %in% c("Bank Day", "Holidays")) %>%
  ggplot(.,
         aes(
           y = values,
           x = abs_difference,
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0, size = 0.3) +
  geom_point(size = 4, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  facet_wrap(~ calendar_variable) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() +
  theme(axis.text.x = element_text(margin = ggplot2::margin(t = 0, unit = "cm")))

# print the plot
graph_love_plot_bank_holidays

# save the plot
ggsave(
  graph_love_plot_bank_holidays,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_love_plot_bank_holidays.pdf"
  ),
  width = 16,
  height = 8,
  units = "cm",
  device = cairo_pdf
)
```

Plot for days of the week:

```{r, fig.width=10, fig.height=5}
# graph for weekdays
graph_love_plot_weekday <- data_calendar %>%
  filter(calendar_variable == "Day of the Week") %>%
  mutate(
    values = fct_relevel(
      values,
      "Monday",
      "Tuesday",
      "Wednesday",
      "Thursday",
      "Friday",
      "Saturday",
      "Sunday"
    )
  ) %>%
  ggplot(.,
         aes(
           y = fct_rev(values),
           x = abs_difference,
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0, size = 0.3) +
  geom_point(size = 4, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() +
  theme(axis.text.x = element_text(margin = ggplot2::margin(t = 0, unit = "cm")))


# print the plot
graph_love_plot_weekday

# save the plot
ggsave(
  graph_love_plot_weekday,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_love_plot_weekday.pdf"
  ),
  width = 16,
  height = 8,
  units = "cm",
  device = cairo_pdf
)
```

Plot for months:

```{r, fig.width=10, fig.height=5}
# graph for month
graph_love_plot_month <- data_calendar %>%
  filter(calendar_variable == "Month") %>%
  mutate(
    values = fct_relevel(
      values,
      "January",
      "February",
      "March",
      "April",
      "May",
      "June",
      "July",
      "August",
      "September",
      "October",
      "November",
      "December"
    )
  ) %>%
  ggplot(.,
         aes(
           y = fct_rev(values),
           x = abs_difference,
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0, size = 0.3) +
  geom_point(size = 4, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  ggtitle("Month") +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() +
  theme(axis.text.x = element_text(margin = ggplot2::margin(t = 0, unit = "cm")))

# print the plot
graph_love_plot_month

# save the plot
ggsave(
  graph_love_plot_month,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_love_plot_month.pdf"
  ),
  width = 16,
  height = 8,
  units = "cm",
  device = cairo_pdf
)
```

Plot for years:

```{r, fig.width=10, fig.height=5}
# graph for year
graph_love_plot_year <- data_calendar %>%
  filter(calendar_variable == "Year") %>%
  ggplot(.,
         aes(
           y = as.factor(as.numeric(values)),
           x = abs_difference,
           colour = fct_rev(dataset),
           shape = fct_rev(dataset)
         )) +
  geom_vline(xintercept = 0, size = 0.3) +
  geom_point(size = 4, alpha = 0.8) +
  scale_colour_manual(name = "Dataset:", values = c(my_blue, my_orange)) +
  scale_shape_manual(name = "Dataset:", values = c(17, 16)) +
  ggtitle("Year") +
  xlab("Absolute Difference in Percentage Points") +
  ylab("") +
  theme_tufte() +
  theme(axis.text.x = element_text(margin = ggplot2::margin(t = 0, unit = "cm")))


# print the graph
graph_love_plot_year

# save the plot
ggsave(
  graph_love_plot_year,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_love_plot_year.pdf"
  ),
  width = 16,
  height = 8,
  units = "cm",
  device = cairo_pdf
)
```

# Overall Balance Improvement

We finally plot the distribution of standardized mean differences for continuous covariates or the absolute percentage points differences for categorical covariates between treated and control units before and after matching.

### Continuous Covariates

```{r, fig.width=6, fig.height=4}
# we select the dataset indicator and the standardized difference
data_love_pollutants <- data_love_pollutants %>%
  ungroup() %>%
  select(dataset, standardized_difference)

data_love_continuous_weather <- data_love_continuous_weather %>%
  ungroup() %>%
  select(dataset, standardized_difference)

data_continuous_love <-
  bind_rows(data_love_pollutants, data_love_continuous_weather)

# create the graph
graph_boxplot_continuous_balance_improvement <-
  ggplot(data_continuous_love,
         aes(x = dataset, y = standardized_difference)) +
  ggbeeswarm::geom_quasirandom(
    shape = 16,
    size = 2,
    width = 0.1,
    color = my_blue,
    alpha = 0.8
  ) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  xlab("Dataset") +
  ylab("Standardized\nMean Differences") +
  ggtitle("Continuous Variables") +
  theme_tufte() +
  theme(axis.text.x = element_text(margin = ggplot2::margin(t = 0, unit = "cm")))

# print the graph
graph_boxplot_continuous_balance_improvement

# save the graph
ggsave(
  graph_boxplot_continuous_balance_improvement,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_boxplot_continuous_balance_improvement.pdf"
  ),
  width = 12,
  height = 8,
  units = "cm",
  device = cairo_pdf
)
```

### Categorical Covariates

```{r, fig.width=6, fig.height=4}
# we select the dataset indicator and the standardized difference
data_calendar <- data_calendar %>%
  ungroup() %>%
  select(dataset, abs_difference)

data_weather_categorical <- data_weather_categorical %>%
  ungroup() %>%
  select(dataset, abs_difference)

data_categorical_love <-
  bind_rows(data_calendar, data_weather_categorical)

# create the graph
graph_boxplot_categorical_balance_improvement <-
  ggplot(data_weather_categorical, aes(x = dataset, y = abs_difference)) +
  ggbeeswarm::geom_quasirandom(
    shape = 16,
    size = 2,
    width = 0.2,
    color = my_blue,
    alpha = 0.8
  ) +
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5)) +
  xlab("Dataset") +
  ylab("Absolute Difference \nin Percentage Points") +
  ggtitle("Categorical Variables") +
  theme_tufte() +
  theme(axis.text.x = element_text(margin = ggplot2::margin(t = 0, unit = "cm")))


# print the graph
graph_boxplot_categorical_balance_improvement

# save the graph
ggsave(
  graph_boxplot_categorical_balance_improvement,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_boxplot_categorical_balance_improvement.pdf"
  ),
  width = 12,
  height = 8,
  units = "cm",
  device = cairo_pdf
)
```

### Combining Plots

We combine the two previous plots:

```{r, fig.width=8, fig.height=4}
# combine the two previous plots
graph_overall_balance  <-
  graph_boxplot_continuous_balance_improvement + graph_boxplot_categorical_balance_improvement +
  plot_annotation(tag_levels = 'A')

# save the graph
ggsave(
  graph_overall_balance,
  filename = here::here(
    "3.outputs",
    "2.matching_analysis",
    "graph_overall_balance.pdf"
  ),
  width = 20,
  height = 10,
  units = "cm",
  device = cairo_pdf
)
```
